Secure Composition of Robust and Optimising Compilers

### 2023-06-20 popl/main.tex
SMS compiler

The only case where the compiler does something interesting is for top-level functions.
Here, a dynamic typecheck is inserted to respect the calling-convention.
Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ calls this function expecting a natural number with $\trg{\lpair{17}{29}}$.
The compiler ensures that execution does not proceed in cases like that.

\MKin{x-lang relations}
\MKin{backtranslation}
\MKin{rtp of tms}

@@

\MKin{compiler $\trg{L}\to\irl{L_{MS}}$}
\MKin{backward sim}
\MKin{rtp of sms}

\MKin{composition: rtp of ms}

@@

\MKin{compiler $\irl{L_{MS}}\to\irl{L_{MS}}$ doing \gls{dce}}
\MKin{compiler $\irl{L_{MS}}\to\irl{L_{MS}}$ doing \gls{cf}}

>>>>

The only case where the compiler does something interesting is for top-level functions.
Here, a dynamic typecheck is inserted to respect the calling-convention.
Since $\trg{L}$ has no static typechecks, it could happen that a bogus context $\trg{\library_{\ctx}}$ calls a symbol accepting a $\src{\natt}$ with $\trg{\lpair{17}{29}}$.
The compiler ensures that execution does not proceed in cases like that.
This paper considers symbols on $\src{\natt}$.

\MKin{x-lang relations $\approx_{\delta;\src{L}}$, $\multimap_{\delta;\src{L}}$, as well as $\cong_{\delta;\src{X}}$}
\MKin{backtranslation}

\begin{theorem}[Compiler $\cca$ is secure with respect to \gls{tms}]\label{thm:cca:rtp:tms}
  $\rtp{\cca}{\tmssafe}$ % \Coqed
\end{theorem}

@@

In this section, the language $\irl{L_{\mssafe}}$ is exactly the same as $\trg{L}$ and are considered equal.
However, it is still named and collored differently to emphasize the fact that after this section, compiled $\irl{L_{\mssafe}}$ programs are \gls{ms}.

\begin{center}
  $$
  \begin{array}{rcl}
    \ccb(\trg{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{1}]})}{\lnew[\irl]{\irl{x}}{\irl{x_{SIZE}}}{\ccb(\trg{\expr[_{2}]})}} \\
    \ccb(\trg{\lget{x}{\expr}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr})}{\irl{\lifz{0\le x_{n}<x_{SIZE}}{\lget{x}{x_{n}}}{\labort}}} \\
    \ccb(\trg{\lset{x}{\expr[_{1}]}{\expr[_{2}]}}) & = & \llet[\irl]{\irl{x_{n}}}{\ccb(\trg{\expr[_{1}]})}{\lifz[\irl]{\irl{0\le x_{n}<x_{SIZE}}}{\lset[\irl]{\irl{x}}{\irl{x_{n}}}{\ccb(\trg{\expr[_{2}]})}}{\irl{\labort}}} \\
  \end{array}
  $$
\end{center}

Up to getting and setting memory, the compiler is the identity function.
For memory accesses, however, a bounds--check is inserted that enforces \gls{sms}.
To this end, the compiler introduces another, fresh identifier $\irl{x_{SIZE}}$ for each allocation that binds $\irl{x}$ to keep track of the allocation size.

\MKin{backward sim}
\begin{theorem}[Compiler $\ccb$ is secure with respect to \gls{sms}]\label{thm:ccb:rtp:sms}
  $\rtp{\ccb}{\smssafe}$ % \Coqed
\end{theorem}

The composition of $\cca$ and $\ccb$ attains full memory safety using \Cref{thm:rtp,thm:cca:rtp:tms,thm:ccb:rtp:sms}.

\begin{theorem}[Compiler $\cca\circ\ccb$ is secure with respect to \gls{ms}]\label{thm:ccab:rtp:ms}
  $\rtp{\cca\circ\ccb}{\mssafe}$ % \Coqed
\end{theorem}

@@

\MKin{compiler $\ccdce$ doing \gls{dce}}
\MKin{compiler $\cccf$ doing \gls{cf}}
